---
title: "Portfilo backtings with Rmetrics"
date: '2017-12-25'
output: 
  html_notebook: 
    fig_caption: yes
    number_sections: yes
    toc: yes
---


```{r setup, include=FALSE  }

library(tidyverse)
library(lubridate)
library(stringr)
library(fPortfolio)
library(zstmodelr)

# Set global options for knitr 
knitr::opts_chunk$set(
comment = "#>",
echo = FALSE,
warning = FALSE,
message = FALSE,
collapse = TRUE
)

options(tibble.print_min = 10L, tibble.print_max = 20L)

```

# Data Preparation(数据准备)

## Data tables in database(数据库所包含数据表)

```{r connect_db}
# Open gta stock databse
stock_db <- stock_db(gta_db, "GTA_SQLData")
open_stock_db(stock_db)

# Initate the stock database
invisible(init_stock_db(stock_db))
stock_name_list  <- stock_name_list(stock_db)
stock_field_list <- stock_field_list(stock_db)

# Get names of all stock table from database   
 db_tables <- list_stock_tables(stock_db)
 knitr::kable(
  x = db_tables,
  caption = "Tables in GTA Stock Database",
  align = "l"
 )
 
```

## Load return timeseris of market and stocks(加载市场与个股回报时间序列)

```{r load_return_series}
# load monthly return of stocks
ds_trd_mnth.df <- get_table_dataset(stock_db, 
                                    table_name = "TRD_Mnth_月个股回报率")
ds_stocks_mretwd.fts <- fetch_stock_field_dataset(ds_source.df = ds_trd_mnth.df,
                                                  stock_cd_list = stock_cd_list, 
                                                  replaceNA = "zeros",
                                                  target_field = "mretwd", 
                                                  date_field = "trdmnt")


# load montly return of market
ds_trd_cnmont.df <- get_table_dataset(stock_db, 
                                    table_name = "TRD_Cnmont_综合市场月度回报")
ds_market_cmretwdtl.fts <- fetch_stock_field_dataset(ds_source.df = ds_trd_cnmont.df,
                                                  stock_cd_list = stock_cd_list, 
                                                  replaceNA = "zeros",
                                                  target_field = "cmretwdtl", 
                                                  date_field = "trdmnt")




```

## Z-score Factor model

### Z-score Stock Selection

筛选最近一期z分值最高的公司如下：

```{r zscore_filter}

# Compute the z-score of indicatros
fundamental_factors_zscore_by_industry <- factors_zscore(fundamental_factors, 
                                             aggregate_formula = ROCE,
                                             group_by = c("date","indcd"))

# Select top stocks basing on z-scores
top_score_stocks_list <- zscore_filter_stocks(fundamental_factors_zscore_by_industry,
                                              ranking_field = "stk_score",
                                              ranking_number =  5)

# Top stocks with highest zscroe in last quarter
date_last_quarter = max(top_score_stocks_list$date)

top_stocks_last_quater <- top_score_stocks_list %>%
    dplyr::filter( date == date_last_quarter) %>%
    dplyr::mutate(stkname = code2name(stock_db, stkcd, type = "stock"),
                 indname = code2name(stock_db, indcd, type = "industry")) %>%
    dplyr::select(date, stkcd, stkname, indcd, indname, stk_score, 
                  everything(), -periodtype)

top_stocks_last_quater


# Top stocks with hightest average zscorse
top_stocks_average_score <- fundamental_factors_zscore_by_industry %>%
    dplyr::ungroup() %>%
    dplyr::group_by(stkcd) %>%
    dplyr::summarise(stk_score = mean(stk_score, na.rm = TRUE), obs = n()) %>%
    dplyr::filter(obs >= 8) %>%
    dplyr::top_n(100, stk_score) %>%
    dplyr::mutate(stkname = code2name(stock_db, stkcd, type = "stock"), 
                  No. = min_rank(desc(stk_score))) %>%
    dplyr::arrange(desc(stk_score)) %>%
    dplyr::select(No., stkcd, stkname, everything())

top_stocks_average_score

```

# Model Exploration(模型探索)



### Z-score Stock Model Paramters Estimation


## Fundenmental factor model



### Estimation of Factor Premium


## Ecomomic Factor Model


### Estimation of Factor Exposure


# Build Portfolio(构建组合)

# Backtest Portfolio(组合回测)
