---
title: "Analysis of Stock Return"
output: 
  html_notebook: 
    fig_caption: yes
    number_sections: yes
    toc: yes
---

研究各股票的投资回报情况


```{r setup, include=FALSE  }

library(fPortfolio)
library(tidyverse)
library(lubridate)
library(stringr)
library(shiny)
library(zstmodelr)

# Set global options for knitr 
knitr::opts_chunk$set(
comment = "#>",
echo = FALSE,
warning = FALSE,
message = FALSE,
collapse = TRUE
)

options(tibble.print_min = 10L, tibble.print_max = 20L)

```
# Data Preparation(数据准备)

## Data tables in database(数据库所包含数据表)
```{r connect_db}
# Open gta stock databse
stock_db <- stock_db(gta_db, "GTA_SQLData")
open_stock_db(stock_db)

# Initate the stock database
invisible(init_stock_db(stock_db))

# Get names of all stock table from database   
 db_tables <- list_stock_tables(stock_db)
 knitr::kable(
  x = db_tables,
  caption = "Tables in GTA Stock Database",
  align = "l"
 )
 
```

**共有`r length(db_tables)`个数据表**

## Summary of Stocks&Industries in Universe(全部股票及行业信息)

**全部股票与行业信息如下：**
```{r summary_stock_info}

# Get company info 
trd_company.df <- get_table_dataset(stock_db, table_name = "TRD_Co_公司基本情况")
trd_company.tib <- tibble::as.tibble(trd_company.df)

# all stocks in the investment universe
universe_stocks <- trd_company.tib %>%
    dplyr::select(stkcd, stkname= stknme, indcd = nnindcd, indname = nnindnme) %>%
    dplyr::arrange(indcd, stkcd)


# all industries in investment universe
universe_industries <- universe_stocks %>%
  dplyr::group_by(indcd, indname) %>%
  dplyr::summarise(stknum = n()) %>%
  dplyr::arrange(indcd, desc(stknum))

# Output the results
universe_industries
universe_stocks

```

## Set Study Params (设置研究参数)

**组合股票明细：**
```{r set_params_study}
# Set params of study

# Set stocks as assets
stock_stkcds_list <- c(600066,000550, 600031, 000157,000651, 000333)

#period_type: c("daily", "weekly", "monthly", "annual")
period_type <- "monthly"

#return_type: c("simple", "compound")
return_type <- "simple"

# start and end date of period
# period_start_date <- today() - 365
period_start_date <- NULL
period_end_date <- today()

#nYearsBack: c("1y", "2y", "3y", "5y", "10y", "YTD")
nYearsBack <- NULL







```
## Load Data of Assets Return (加载资产投资回报数据)
```{r get_assets_data}

# Get stocks and market return data
stocks_return <- get_stock_return(stock_db, 
                                  stock_stkcds_list, 
                                  period_type = period_type, 
                                  return_type = return_type,
                                  use_stock_name = FALSE,
                                  output_type = "timeSeries")

market_return <- get_market_return(stock_db, 
                                   period_type = period_type, 
                                   return_type = return_type,
                                   output_type = "timeSeries")

# Build assets_return
assets_return <- get_assets_return(benchmark_return = market_return,
                                         stocks_return = stocks_return)
if (!is.null(nYearsBack)) {
  # get lastest series
  assets_return <- endOfPeriodSeries(assets_return)
  
} else {
  # get window series
  if ( !is.null(period_start_date) && !is.null(period_end_date)) {
    assets_return <- window(assets_return, start = period_start_date,
                             end = period_end_date)
  }
}
                    
assets_cumulated_return <- cumulated_return(assets_return, method = "simple")

```


# Data Exploration(数据探索)
## Basic Statistics of Assets Return(资产回报基本统计信息分析)
```{r assets_return_summary}

assets_return_stats <- basicStats(assets_return)
print(round(assets_return_stats, digits = 3))

# display assets box plot
assetsBoxPlot(assets_return,cex = 0.4)
 
# display assets box percentile plot
assetsBoxPercentilePlot(assets_return)



```



## Comparing Multivariate Assets Return Statistics(资产回报统计对比)
```{r assets_return_compare_multivariate}

# Segment Plots of Box Plot Statistics
assetsBoxStatsPlot(assets_return)

# Segment plots of basic return statistics
assetsBasicStatsPlot(assets_return)

# Segment Plots of Distribution Moments
assetsMomentsPlot(removeNA(assets_return))


```


## Pairwise Dependencies of Assets Return(资产回报相关性分析)
```{r assets_return_pairwise_dependence}

histPanel <- function(x, ...) {
  usr <- par("usr")
  on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1.5))
  h <- hist(x, plot = FALSE)
  breaks <- h$breaks
  nB <- length(breaks)
  y <- h$counts
  y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, ...)
}

# Simple Pairwise Scatter Plots of Assets
assetsPairsPlot(assets_return, 
                main = "Simple Pairwise Scatter Plots of Assets", 
                pch = 19, cex = 0.5, col = "royalblue4",
                 diag.panel = histPanel)

# Pairwise Correlations Between Assets
assetsCorgramPlot(assets_return, method = "pie", 
                  main = "Pairwise Correlations Between Assets", 
                  pch = 19, cex = .5, 
                  diag.panel = histPanel)
assetsCorgramPlot(assets_return, method = "shade", 
                  main = "Pairwise Correlations Between Assets", 
                  pch = 19, cex = .5, 
                  diag.panel = histPanel)

# Tests of Pairwise Correlations
assetsCorTestPlot(removeNA(assets_return), 
                  main = "Tests of Pairwise Correlations", 
                  diag.panel = histPanel)

# Image Plot of Correlations
assetsCorImagePlot(assets_return)



```


## Modelling Asset Returns(资产回报建模分析)
```{r asset_return_modelling}

# display asset HistPlot
for (i in 1:ncol(assets_return)) {
  assetsHistPlot(removeNA(assets_return[, i]))
}

# display normal qq plot 
for (i in 1:ncol(assets_return)) {
  assets_stkcd <- names(assets_return)[i]
  if(assets_stkcd != "market_index"){
    assets_name <- code2name(stock_db, as.numeric(assets_stkcd))
  } else {
    assets_name <- ""
  }
  sub_titile <- sprintf("%6s:%s", assets_stkcd, assets_name )
  assetsQQNormPlot(removeNA(assets_return[, i]), sub = sub_titile)
}

```



## Drawdown Statistics of Assets Return(资产回报回撤分析)
```{r assets_return_drawndowns}

# display drawndown and turns plot 
for (i in 1:ncol(assets_return)) {

  drawdownsAnalytics(assets_cumulated_return[, i], trace = FALSE)
  
  turnsAnalytics(assets_cumulated_return[, i], trace = TRUE)
  
  asset_drawdown_stats <- drawdownsStats(removeNA(assets_return[, i]))
  print(asset_drawdown_stats, digits = 2)
  
}


```





## Group of Similar or Dissimlar Assets(类似资产分组)

**A.Grouping Asset Returns by Hierarchical Clustering**
```{r assets_grouping_hclust}

assets_group_hclust_complete <- assetsSelect(assets_return, method = "hclust")
plot(assets_group_hclust_complete, xlab = "Assets")
mtext("Distance Metric: Euclidean", side = 3)

assets_group_hclust_Ward <- assetsSelect(assets_return, method = "hclust", 
                                   control = c(measure = "euclidean",
                                               method = "ward"))

plot(assets_group_hclust_Ward, xlab = "Assets")
mtext("Distance Metric: Euclidean", side = 3)

```

**B.Grouping Asset Returns by k-means Clustering**
```{r assets_grouping_kmeans}

assets_group_kmeans <- assetsSelect(removeNA(assets_return), method = "kmeans", 
                              control <- c(centers = 3,
                                           algorithm = "Hartigan-Wong"))
sort(assets_group_kmeans$cluster)

```


**C.Grouping Asset Returns through Eigenvalue Analysis**
```{r assets_grouping_pca}

assets_group_pca_pearson <- assetsCorEigenPlot(assets_return, method = "pearson")
mtext("Method: pearson", side = 3)

assets_group_pca_kendall <- assetsCorEigenPlot(assets_return, method = "kendall")
mtext("Method: kendall", side = 3)

assets_group_pca_spearman <- assetsCorEigenPlot(assets_return, method = "spearman")
mtext("Method: spearman", side = 3)

```

**D.Grouping Asset Returns by Contributed Cluster Algorithms**
```{r assets_grouping_other_cluster}
.pamSelect <- function(x, control = NULL, ...) {
    if (is.null(control))
    control = c(k = 2, metric = "euclidean")
    k <- as.integer(control[1])
    metric <- control[2]
    cluster::pam(x <- as.matrix(x), k = k, metric = metric, ...)
}

assets_group_pam <- assetsSelect( removeNA(assets_return), 
                                  method = "pam", 
                                  control <- c(k = 2, metric = "euclidean"))

plot(assets_group_pam, which.plots = 1)

```


# Portfolio investment Strategy(投资组合策略)
## Basic Portfolios(基本投资组合)

```{r portfolio_setting}

portfolio_bemchmark_return <- removeNA(assets_return)
portfolio_return <- portfolio_bemchmark_return[, -1]

portfolio_assets_name <- names(portfolio_return)
portfolio_assets_code <- name2code(stock_db, portfolio_assets_name, type = "stock")

```
**投资组合中的股票资产**

资产代码：`r portfolio_assets_code`

资产代码：`r portfolio_assets_name`

共有`r nrow(portfolio_return)`个数据记录




### Mean-Variance Portfolios

**A.Compute Equal Weights Portfolio**
```{r mv_portfolio_equalweights}

equalwieght_portfolio_spec <- portfolioSpec()
setWeights(equalwieght_portfolio_spec) <- rep(1/ncol(portfolio_return), 
                                              times = ncol(portfolio_return))
equalweight_portfolio <- feasiblePortfolio( data = portfolio_return,
                                            spec = equalwieght_portfolio_spec,
                                            constraints = "LongOnly" 
                                            )
print(equalweight_portfolio)

weightsPie(equalweight_portfolio, radius = 0.7)
mtext(text = "Equally Weighted MV Portfolio: weights", side = 3, line = 1.5,
font = 2, cex = 0.7, adj = 0)

weightedReturnsPie(equalweight_portfolio, radius = 0.7)
mtext(text = "Equally Weighted MV Portfolio: weightedReturns", side = 3, line = 1.5,
font = 2, cex = 0.7, adj = 0)

covRiskBudgetsPie(equalweight_portfolio, radius = 0.7)
mtext(text = "Equally Weighted MV Portfolio: covRiskBudgets", side = 3, line = 1.5,
font = 2, cex = 0.7, adj = 0)

```


**B.Compute a Mininum Risk Efficient Portfolio**
```{r mv_porfolio_minRisk}

minrisk_portfolio_spec <- portfolioSpec()
setTargetReturn(minrisk_portfolio_spec) <- getTargetReturn(equalweight_portfolio)["mean"]
minrisk_portfolio <- efficientPortfolio( data = portfolio_return,
                                            spec = minrisk_portfolio_spec,
                                            constraints = "LongOnly" 
                                            )
print(minrisk_portfolio)

weightsPie(minrisk_portfolio, radius = 0.7)
mtext(text = "Minimal Risk MV Portfolio: weights", side = 3, line = 1.5,
font = 2, cex = 0.7, adj = 0)

weightedReturnsPie(minrisk_portfolio, radius = 0.7)
mtext(text = "Minimal Risk MV Portfolio: weightedReturns", side = 3, line = 1.5,
font = 2, cex = 0.7, adj = 0)

covRiskBudgetsPie(minrisk_portfolio, radius = 0.7)
mtext(text = "Minimal Risk MV Portfolio: covRiskBudgets", side = 3, line = 1.5,
font = 2, cex = 0.7, adj = 0)

```


**C.Compute the Global Minimum Variance Portfolio**
```{r mv_portfolio_glboalMinVar}

globalMinVar_portfolio_spec <- portfolioSpec()

globalMinVar_portfolio <- minvariancePortfolio( data = portfolio_return,
                                            spec =globalMinVar_portfolio_spec,
                                            constraints = "LongOnly" 
                                            )
print(globalMinVar_portfolio)

weightsPie(globalMinVar_portfolio, radius = 0.7)
mtext(text = "Global Minimum Variance MV Portfolio: weights", side = 3, line = 1.5, 
      font = 2, cex = 0.7, adj = 0)

weightedReturnsPie(globalMinVar_portfolio, radius = 0.7)
mtext(text = "Global Minimum Variance MV Portfolio: weightedReturns", side = 3, line = 1.5,
font = 2, cex = 0.7, adj = 0)

covRiskBudgetsPie(globalMinVar_portfolio, radius = 0.7)
mtext(text = "Global Minimum Variance MV Portfolio: covRiskBudgets", side = 3, line = 1.5,
font = 2, cex = 0.7, adj = 0)

```


**D.Compute the Tangency Portfolio**
```{r mv_portfolio_tangency}

tangency_portfolio_spec <- portfolioSpec()
setRiskFreeRate(tangency_portfolio_spec) <- 0.00
tangency_portfolio <- tangencyPortfolio( data = portfolio_return,
                                            spec = tangency_portfolio_spec,
                                            constraints = "LongOnly" 
                                            )
print(tangency_portfolio)

weightsPie(tangency_portfolio, radius = 0.7)
mtext(text = "Tangency MV Portfolio: weights", side = 3, line = 1.5,
font = 2, cex = 0.7, adj = 0)

weightedReturnsPie(tangency_portfolio, radius = 0.7)
mtext(text = "Tangency MV Portfolio: weightedReturns", side = 3, line = 1.5,
font = 2, cex = 0.7, adj = 0)

covRiskBudgetsPie(tangency_portfolio, radius = 0.7)
mtext(text = "Tangency MV Portfolio: covRiskBudgets", side = 3, line = 1.5,
font = 2, cex = 0.7, adj = 0)

```

**E.MV Portfolio Frontier Plot**
```{r mv_portfolio_Frontier}

long_portfolio_spec <- portfolioSpec()

long_portfolio_frontier <- portfolioFrontier(data = portfolio_return, 
                                             spec = long_portfolio_spec)
tailoredFrontierPlot(object = long_portfolio_frontier, 
                     mText = "MV Portfolio - LongOnly Constraints",
                     risk = "Cov")

# create weights related plots

weightsPlot(long_portfolio_frontier, mtext = FALSE)
# text <- "Mean-Variance Portfolio - Long Only Constraints"
# mtext(text, side = 3, line = 3, font = 2, cex = 0.9)
weightedReturnsPlot(long_portfolio_frontier, mtext = FALSE)
covRiskBudgetsPlot(long_portfolio_frontier, mtext = FALSE)


```



## Portfolio Backtesting(投资策略回测)

回测期间：`r start(portfolio_bemchmark_return)` -> `r end(portfolio_bemchmark_return)`


```{r backtest_strategy}
  
my_portfolio_strategy <- function(data, spec, constraints, backtest)
{
    ## Extract Parameters:
    Parameters <- getStrategyParams(backtest)
    ## Strategy Portfolio:
    strategyPortfolio <- tangencyPortfolio(data, spec, constraints)
    ## Return :
    strategyPortfolio
}

portfolio_spec <- portfolioSpec()
portfolio_backtest <- portfolioBacktest()
setStrategyFun(portfolio_backtest) <- "tangencyStrategy"
# setStrategyFun(portfolio_backtest) <- "my_portfolio_strategy"
setStrategyParams(portfolio_backtest) <- list()

portfolio_backtest_fomula <- market_index ~ I(宇通客车) + I(江铃汽车) + I(三一重工) + I(中联重科) + I(格力电器) + I(美的集团)
# portfolio_backtest_fomula <- market_index ~ 600066 + 000550 + 600031 + 000157 + 000651 + 000333
# portfolio_backtest_fomula <- market_index ~ "600066"+"000550"


# portfolio_backtest_fomula <- market_index ~ market_index

portfolio_constraints <- "LongOnly"


backtest_portfolios <- portfolioBacktesting(formula = portfolio_backtest_fomula,
                                      data = portfolio_bemchmark_return, 
                                      spec = portfolio_spec,
                                      constraints = portfolio_constraints,
                                      backtest = portfolio_backtest,
                                      trace = FALSE)

backtest_soomthing_portfolios <- portfolioSmoothing(object = backtest_portfolios,                                                                               trace = FALSE)

backtestPlot(backtest_soomthing_portfolios, 
             cex = 0.6, font = 1, family = "mono", 
             format = "%Y-%m")

netPerformance(backtest_soomthing_portfolios)

```

